---
title: "Markov Model Attribution"
author: "Sean Astrup - seanastrup@gmail.com"
runtime: shiny
css: www/my.css
output: 
  html_notebook:
    code_folding: hide
    theme: united
---

####Post in progress. Stable development release updated on: 2018-10-20
Current State:   
* Conversion data is **randomly generated for development and testing**. Results should **NOT** currently be used for **ANYTHING**. 

##Markov multi-channel conversion optimization
I may do a follow-up or expansion to this on *why* multi-channel funnel optimization is important as well as a deeper explanation of the methods and models used. For now, this tool may be useful to those analysts or campaign managers with a general understanding of *at least* what markov chains are as well as their place in an overall multi-channel optimization strategy.

In order to use this tool you'll need:  

1. Access to a Google Analytics account with multi-channel funnels enabled and set up:
    i) We need to set up a custom, destination-based goal on each Google Analytics view in order to measure visits through multi-channel funnels so if you are an agency without access to create goals on client accounts, you'll need to coach them through that.
    ii) Once you have access to create goals, create a new custom goal that will count a goal completion **every time a user visits**. If you don't know how to set this up, follow my separate post with instructions here [post link]

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(DT)
library(googleAnalyticsR)
library(googleAuthR)
library(ChannelAttribution)
library(highcharter)
library(ggplot2)
source('ga-auth.R')
DEPLOYMENT_TAG <- 'public'
```

```{r ga_auth, echo=FALSE}
googleAuthUI("auth")
token  <- callModule(googleAuth, "auth")

###################################################
########## VIEW ID SELECTION CONTROLS #############
###################################################
  ga_accounts <- reactive({
    
    req(token())
    
    with_shiny(ga_account_list, shiny_access_token = token())
    
  })
  
  # selected_id <- callModule(authDropdown, 'auth_menu', ga.table = ga_accounts)
  
  AccountTable <- reactive({
    
    ga_accounts()
    
  })
  
  pList <- reactive({
    
    ga.table <- ga_accounts()
    
    ga.table[,c('accountName','webPropertyId','websiteUrl','viewName', 'viewId')]
    
  })
  
  ##############################
  ### UPDATE SELECTION VALUES ##
  ##############################
  observe({
    
    validate(
      need(pList(), "Need profiles")
    )
    
    pList  <- pList()
    
    AccountChoices <- unique(pList$accountName)
    
    updateSelectizeInput(session, 
                      "Accounts",
                      label = "Accounts",
                      choices = AccountChoices, 
                      selected = AccountChoices[1],
                      server = TRUE)
    
  })
  
  observe({
      
      validate(
        need(input$Accounts, "Need accounts")
      )
      pList  <- pList()
      
      pList <- pList[input$Accounts == pList$accountName,]
      
      PropertyChoices <- pList$websiteUrl
      
      updateSelectizeInput(session, 
                        "Properties", label="Properties",
                        choices = PropertyChoices, 
                        selected = PropertyChoices[1],
                        server = TRUE)
    
    })
  
  observe({
    
    validate(
      need(input$Properties, "Need web")
    )
    
    pList <- pList()
    
    pList <- pList[input$Properties == pList$websiteUrl,]
    
    ViewChoices <- pList$viewId 
    
    names(ViewChoices) <- paste(pList$viewName, pList$viewId)
    
    updateSelectizeInput(session, 
                         "Views", label = "Views",
                         choices = ViewChoices, 
                         selected = ViewChoices[1],
                         server = TRUE)
  })
  
  GaAccount <- reactive({
    
    validate(
      need(pList(), "Need profiles")
    )
    
    pList  <- pList()
    
    out <- unique(pList$accountName)
    
    return(out[1])
    
  })
  GaProperty <- reactive({
    
    validate(
      need(input$Properties, "Please login")
    )
    
    pList <- pList()
    out <- pList[input$Accounts == pList$accountName,]
    
    return(out$websiteUrl)
    
  })
  GaView <- reactive({
    
    validate(
      need(input$Views, "Please login")
    )
    pList <- pList()
    
    out <- pList[input$Views == pList$viewId,]
    
    return(out$viewId)
    
  })
  GaGoal <- reactive({
    
    validate(
      need(input$GoalId, 'Choose Goal Id')
    )
    
    return(input$GoalId)
    
  })
  
```
```{r}
shiny::radioButtons(inputId = 'TestSwitch', label = '', 
                    inline = TRUE, 
                    choices = c('Example Post Data' = 'TestData',
                                'Use My Own Data' = 'UserData')
                    )
```

```{r echo=FALSE}
shiny::inputPanel(
  shiny::selectizeInput(inputId = 'Accounts', label = 'Accounts', choices = NULL, 
                        width = '90%'),
  shiny::selectizeInput(inputId = 'Properties', label = 'Properties', choices = NULL, width = '90%'),
  shiny::selectizeInput(inputId = 'Views', label = 'Views', choices = NULL, 
                        width = '90%'), 
  shiny::textInput(inputId = 'GoalId', label = 'Every Visit Goal Id',
                   placeholder = '3 digit custom goal ID', width = '90%')
) 
shiny::radioButtons(inputId = 'DateFilter', label = 'Dates', inline = TRUE,
                    choices = c('1 week' = 'oneWeek', '1 month' = 'oneMonth', '3 months' = 'threeMonths', 
                                '1 year' = 'oneYear'))
```

&nbsp;  

## Raw Multi-Channel Path Information
We start with taking a look at what our raw data looks like. A simple pull from the multi-channel report section of Google Analytics is a very powerful piece of information. ***Note:*** when viewing non-test data, some cleaning has been done to the source/medium paths (e.g. removing spaces and spam): 
```{r echo = FALSE}
renderDataTable({
  
  Data <- test_data()[, 2:4]
  
  datatable(Data,
            rownames = FALSE,
            options = list(
              pageLength = 5, 
              searching = FALSE, 
              bLengthChange = FALSE
  ))

})
```
&nbsp;  

Each record here represents a particular source/medium path combination and its result in either a purchase (transactions), or non-converted session (null). If you've enabled multi-channel funnels in Google Analytics and have also set up a custom goal that records a completion every time someone visits your site, you can enter that goal ID above to see the **overall conversion rate** across your source medium paths:

```{r echo = FALSE}
renderDataTable({
  
  Data <- make_total_data()
  
  datatable(Data, 
            rownames = FALSE,
            options = list(
              pageLength = 5, 
              searching = FALSE, 
              bLengthChange = FALSE
            ))
  
})
```
&nbsp;  

## Modeling The Information  
Using the handy ChannelAttribution R package, we can evalueate each source/medium and assign a total conversion number to each as well as conversion numbers using first, last, and linear touch models.

```{r echo = FALSE}
renderDataTable({
  
  Data <- ModeledData()
  
  datatable(Data, 
            rownames = FALSE, 
            options = list(
              pageLength = 5, 
              searching = FALSE, 
              bLengthChange = FALSE))
  
})
```
&nbsp;  

## Visualizing The Attribution Strategies
A nice feature of the markov model returned from the ChannelAttribution package is the transition matrix. This information indicates the probabliity that a user in a given source/medium will end up in the next (e.g. A user visits the site from "channel_0", and returns to the site through "channel_1"). With the raw data from our model we can make an informative visualization using ggplot2: 
```{r echo = FALSE}
renderPlot({
  
  df_plot_trans <- MarkovModel()$transition_matrix
 
  cols <- c("#e7f0fa", "#c9e2f6", "#95cbee", "#0099dc", "#4ab04a", "#ffd73e", "#eec73a",
            "#e29421", "#e29421", "#f05336", "#ce472e")
  t <- max(df_plot_trans$transition_probability)

  return(
    ggplot(df_plot_trans, aes(y = channel_from, x = channel_to, fill = transition_probability)) +
            theme_minimal() +
            geom_tile(colour = "white", width = .9, height = .9) +
            scale_fill_gradientn(colours = cols, limits = c(0, t),
                                 breaks = seq(0, t, by = t/4),
                                 labels = c("0", round(t/4*1, 2), round(t/4*2, 2), round(t/4*3, 2), round(t/4*4, 2)),
                                 guide = guide_colourbar(ticks = T, nbin = 50, barheight = .5, label = T, barwidth = 10)) +
            geom_text(aes(label = round(transition_probability, 2)), fontface = "bold", size = 4) +
            theme(legend.position = 'bottom',
                  legend.direction = "horizontal",
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  plot.title = element_text(size = 20, face = "bold", vjust = 2, color = 'black', lineheight = 0.8),
                  axis.title.x = element_text(size = 24, face = "bold"),
                  axis.title.y = element_text(size = 24, face = "bold"),
                  axis.text.y = element_text(size = 8, face = "bold", color = 'black'),
                  axis.text.x = element_text(size = 8, angle = 90, hjust = 0.5, vjust = 0.5, face = "plain")) +
            ggtitle("Transition matrix heatmap")
  )
})
```


```{r echo=FALSE}

test_data <- function(){
  
  df2 <- data.frame(client_id = sample(c(1:1000), 5000, replace = TRUE),
                    date = sample(c(1:32), 5000, replace = TRUE),
                    channel = sample(c(0:9), 5000, replace = TRUE,
                                     prob = c(0.1, 0.15, 0.05, 0.07, 0.11, 0.07, 0.13, 0.1, 0.06, 0.16)))
  df2$date <- as.Date(df2$date, origin = "2015-01-01")
  df2$channel <- paste0('channel_', df2$channel)
   
  # aggregating channels to the paths for each customer
  df2 <- df2 %>%
          arrange(client_id, date) %>%
          group_by(client_id) %>%
          summarise(SourceMediumPath = paste(channel, collapse = ' > '),
                    # assume that all paths were finished with conversion
                    transactions = 1,
                    null = 0) %>%
          ungroup()
  
  return(df2)
    
}

test_rate_data <- function(){
  
  Data <- data.frame(SourceMediumPath = c('channel_0', 'channel_1', 'channel_2', 
                                      'channel_3', 'channel_4', 'channel_5', 
                                      'channel_6', 'channel_7', 'channel_8', 
                                      'channel_9'), 
                     Sessions = sample(5000:8000, size = 10), 
                     transactions = sample(500:1000, size = 10))
  
  Data <- Data %>% mutate(ConversionRate = transactions / Sessions)
  
  
  return(Data)
  
}

DateFilter <- reactive ({
  
  switch(input$DateFilter,
   'oneWeek'     = 7,
   'oneMonth'    = 30,
   'threeMonths' = 90,
   'oneYear'     = 365)

})  

get_total_sessions <- reactive({
  
  req(GaView())
  req(GaGoal())
  gaid <- GaView()
  GoalId = GaGoal()
  McfFilter <- paste0('mcf:conversionGoalNumber==', GoalId)
  
  
  Sessions <- 
    with_shiny(google_analytics_3,
               id = gaid,
               start = lubridate::today() - DateFilter(), 
               end =  lubridate::today(),
               metrics = c('totalConversions'),
               dimensions = c('sourceMediumPath'),
               type = 'mcf',
               filters = McfFilter, 
               shiny_access_token = token()
               )
  names(Sessions) <- c('SourceMediumPath','sessions')
  
  # Clean junk from results 
  Sessions$SourceMediumPath <- gsub(" / ", "/", Sessions[,1])              
  Sessions$SourceMediumPath <- gsub(":?(NA|CLICK|NA):?", "", Sessions[,1]) 
  Sessions$sessions <- as.numeric(Sessions[,2]) 
  
  # Sessions$sessions <- sample(3000, size = nrow(Sessions), replace = TRUE)     # random noise for dev
  
  return(Sessions)
  
})

get_total_conversions <- reactive({
  
  req(GaView())
  gaid <- GaView()
  Conversions <- 
    with_shiny(google_analytics_3,
               id = gaid,
               start = lubridate::today() - DateFilter(), 
               end =  lubridate::today(),
               metrics = c('totalConversions'),
               dimensions = c('sourceMediumPath'),
               type = 'mcf',
               filters = 'mcf:conversionType==Transaction',
               shiny_access_token = token()
               )
    
  names(Conversions) <- c('SourceMediumPath', 'transactions')
  
  # Clean junk from results 
  Conversions$SourceMediumPath <- gsub(" / ", "/", Conversions[,1])              
  Conversions$SourceMediumPath <- gsub(":?(NA|CLICK|NA):?", "", Conversions[,1]) 
  Conversions$transactions <- as.numeric(Conversions[,2]) 
  Conversions$transactions <- Conversions$transactions %>% replace_na(0)
    
  Conversions$transactions <- sample(100, size = nrow(Conversions), replace = TRUE)  # random noise for dev
  
  return(Conversions)
  
})

make_total_data <- reactive({
  
  DataPath <- input$TestSwitch
  
  if(DataPath == 'TestData'){
    
    test_data <- test_rate_data()
    return(test_data)
    
  } else{
    
    Conversions <- get_total_conversions()
    Conversions$SourceMediumPath <- as.character(Conversions$SourceMediumPath)
    Sessions <- get_total_sessions()
    Sessions$SourceMediumPath <- as.character(Sessions$SourceMediumPath)
  
    Data <- full_join(Conversions, Sessions, by = 'SourceMediumPath')
    Data$transactions <- as.numeric(Data$transactions)
  
  
    Data <-
      Data %>%
      mutate(ConversionRate = ifelse(transactions == 0, 0, transactions / sessions),
             null = sessions - transactions)
    
    return(Data)
    
  } 
  
})

make_model_data <- reactive({
  
  DataPath <- input$TestSwitch
  
  if(DataPath == 'TestData'){
    
    test_data <- test_data()
    return(test_data)
    
  } else{
    
    Conversions <- get_total_conversions()
    Conversions$SourceMediumPath <- as.character(Conversions$SourceMediumPath)
    Sessions <- get_total_sessions()
    Sessions$SourceMediumPath <- as.character(Sessions$SourceMediumPath)
  
    Data <- full_join(Conversions, Sessions, by = 'SourceMediumPath')
    Data$transactions <- as.numeric(Data$transactions)
  
  
    Data <-
      Data %>%
      mutate(ConversionRate = ifelse(transactions == 0, 0, transactions / sessions),
             null = sessions - transactions)
    
    return(Data)
    
  } 
  
})

MarkovModel <- reactive({
  
  Data <- make_model_data()
  
  Model <- ChannelAttribution::markov_model(Data, var_path = 'SourceMediumPath', 
                                            var_conv = 'transactions', 
                                            var_null = 'null', 
                                            out_more = TRUE)

  return(Model)
  
})

HeuristicModel <- reactive({
  
  Data <- make_model_data()
  
  Model <- ChannelAttribution::heuristic_models(Data, var_path = 'SourceMediumPath',
                                                var_conv = 'transactions')

  return(Model)
  
})

ModeledData <- reactive({
  
  MarkovModel <- MarkovModel()$result
  HeuristicModel <- HeuristicModel()
  
  ModeledData <- full_join(MarkovModel, HeuristicModel, by = 'channel_name')
  
})

```

